### 2024-04-13 docs/SOSP 2024/abstract.tex
Update abstract

  The rapid proliferation of machine learning coupled with accelerating evolution of the hardware ecosystem has led to a surge in the demand for model inference on a variety of hardware.
  %While CPUs have been the mainstay for machine learning inference, the availability of GPUs holds promise to scale to bigger and more powerful models. 
  This paper is motivated by the problems encountered when targeting inference of decision tree based models, the most popular models on tabular data, to run at peak performance on 
  diverse CPU and GPU targets. We evaluated existing solutions and found that they do not provide portable performance across different hardware targets.
  %Decision tree based models are widely used in practice due to their robustness, interpretability, and ability to handle missing data.

  To address this we present the design of \Treebeard{}, a schedule guided compiler  
  for decision tree based models that searches over a large design space 
  to automatically generate high-performance inference routines.
  %We re-architect the open-source \TreebeardOLD{} infrastructure and significantly extend it to enable high-performance code generation across target processors. 
  \Treebeard{} has two core components. A scheduling language 
  that encapsulates the large optimization space for decision tree inference, 
  and techniques to efficiently explore this space.
  \TODO{Change large optimization space}.
  %We also design a set of heuristics that can find near-optimal solutions in quick time.
  Second, an optimizing multi-level compiler that can generate code based on the schedule. 
  For the latter, we re-architect the open-source \TreebeardOLD{} CPU compiler to support schedule guided compilation and
  add support for GPU code generation. GPU code generation required fundamental new optimization interfaces for 
  caching, parallel reduction, and support for multiple in-memory representations of trees.

  %The compiler builds on an existing compiler for CPUs, \TreebeardOLD{}, also add support for caching, parallel reduction and a plug-in mechanism to explore different in-memory representations of trees.

  We evaluate \Treebeard{} on over seven hundred diverse models and demonstrate that the best schedule varies drastically with model, batch size, and target hardware. 
  Our scheduling heuristic is able to quickly find near optimal schedules while searching over a small number (\~50) of schedules.
  In terms of performance, \Treebeard{} generated code is an order of magnitude faster than XGBoost and
  about 2-3$\times$ faster on average than RAPIDs FIL and Tahoe. While these systems only target NVIDIA GPUs, \Treebeard{} achieves competent performance on AMD GPUs as well. 
  On CPUs, \Treebeard{} achieves better scaling compared to \TreebeardOLD{}.
  For models where \TreebeardOLD{} was only able to achieve diminishing returns with an 
  increasing number of threads, \Treebeard{} is able to scale linearly with the number of threads.
  \TODO{(numbers for CPU performance?)}
\end{abstract}

>>>>

  This paper is motivated by the growing demand for the increased performance
  of machine learning applications on different hardware platforms
  including CPUs and GPUs. We focus on accelerating the inference of decision tree based models, 
  which are the most popular models for tabular data. Existing solutions do not achieve 
  the highest possible performance because they do not explore different optimization
  configurations. And since these systems are hand-written, they are not portable either.

  We address these problems by designing \Treebeard{}, a \emph{schedule-guided},
  \emph{retargetable} compiler infrastructure for decision tree based models.  
  \Treebeard{} has two core components. The first is a scheduling language 
  that encapsulates the large optimization space for decision tree inference, 
  and techniques to efficiently explore this space. \TODO{Change large optimization space}.
  The second is an optimizing retargetable multi-level compiler that 
  can generate code for any specified schedule. \Treebeard{}'s retargetability 
  is based not only on being able to generate code for different target architectures (CPU vs. GPU),
  but also on its ability to use different data layouts, caching strategies,
  parallel reduction schemes etc. To accomplish this level of configurability, we re-architect 
  and significantly extend the open-source \TreebeardOLD{} CPU compiler
  to support (i) schedule-guided compilation, (ii) retargetable GPU code
  generation, and (iii) GPU-specific optimizations. 
  %The scheduler and retargetable code generator components work hand-in-glove, synergizing and mutually benefiting from each other.

  We demonstrate that \Treebeard{} can generate high-performance inference 
  code, for several hundred decision tree models across different batch sizes
  and target architectures. Our scheduling heuristic is able to quickly find near-optimal schedules 
  \TODO{[how do we argue that the schedule is near-optimal? Do we have some 
  exptl. evidence that we can show in the results section?]}
  schedule while searching over a small number ($\sim$50) of schedules.
  In terms of performance, \Treebeard{} generated code is an order of magnitude faster than XGBoost and
  about 2-3$\times$ faster on average than RAPIDs FIL and Tahoe. While these systems only 
  target NVIDIA GPUs, \Treebeard{} achieves competent performance on AMD GPUs as well. 
  On CPUs, \Treebeard{} achieves better scaling compared to \TreebeardOLD{}.
  For models where \TreebeardOLD{} was only able to achieve diminishing returns with an 
  increasing number of threads, \Treebeard{} is able to scale linearly with the number of threads.
  \TODO{(numbers for CPU performance?)}
\end{abstract}
